Trajectory Mapping Results

### 2020-12-30 scripts/analyze_trajectories_v2.Rmd
wip dataset trajectories

case_data <- get_cases(demes, from = ymd("2019-09-01"), to = MRS + 10) %>%
  arrange(date) %>%
  group_by(deme) %>%
  mutate(cumcases = cumsum(cases),
         cumdeaths = cumsum(deaths)) %>%
  group_by(deme, date) %>%
  mutate(cumcases = max(cumcases),
         cumdeaths = max(cumdeaths)) %>%
  select(deme, date, cumcases, cumdeaths, cases)  %>%
  distinct_at(vars(deme, date, cumcases, cumdeaths), .keep_all = TRUE)

@@

case_data_pop <- case_data %>% left_join(population, by="deme") %>% mutate(cas100 = cumcases * 1e5 / (pop * 1e3))
```

@@

events <- df$events %>%
  mutate(src = factor(src,
                      levels = 0:(nrow(demes) - 1),
                      #labels = sort(demes$deme)),
                      labels = c("France", "Germany","China","Italy","OtherEuropean","Spain")),
         dest = ifelse(is.na(dest), as.character(src),
                       as.character(factor(dest,
                                           levels = 0:(nrow(demes) - 1),
                                           labels = sort(demes$deme)))),
         date_model = date(date_decimal(decimal_date(ymd(MRS)) - age))) %>%
  mutate(date = date_model + 10)
```

@@

```{r traj-process, warning=FALSE}
# Date grids
max_age <- max(states$age)
min_date <- date_decimal(decimal_date(mrs) - max_age)
#ages_5day <- rev(decimal_date(mrs) - decimal_date(seq(mrs,min_date, by = '-5 days')))
ages_1day <- rev(decimal_date(mrs) - decimal_date(seq(mrs,min_date, by = '-1 days')))

# Grid trajectories, time grid of 5 days taking maximum value
gt <- gridTrajectoriesByAge(states, ages_1day) %>%
  replace_na(list(N = 0)) %>%
  mutate(date_model = date(date_decimal(decimal_date(mrs) - age)),
         date = date_model + 10)

# Summary statistics
gt_summary <- gt %>%
  group_by(date, type) %>%
  summarize(Imean = mean(N),
            Imedian = median(N),
            Ilow    = quantile(N, 0.025), #HPD??
            Ihigh   = quantile(N, 0.975),
            .groups = "drop_last") %>%
  ungroup()

# Grid events, time grid of 1 day, taking the sum of events
ge <- gridEventsByAge(events, ages_1day) %>%
  mutate(date_model = date(date_decimal(decimal_date(mrs) - age)),
         date = date_model + 10,
         dest = ifelse(is.na(dest), as.character(src), as.character(dest)))

@@

>>>>

case_data <- get_cases(demes, from = ymd("2019-09-01"), to = MRS + 10) %>%
  arrange(date) %>%
  group_by(deme) %>%
  rename(value = cases) %>%
  mutate(cumvalue = cumsum(value)) %>%
  group_by(deme, date) %>%
  mutate(cumvalue = max(cumvalue)) %>%
  select(deme, date, cumvalue, value)  %>%
  distinct_at(vars(deme, date, cumvalue), .keep_all = TRUE)

@@

case_data_pop <- case_data %>% left_join(population, by="deme") %>% mutate(ECDCcas100 = cumcases * 1e5 / (pop * 1e3)) 
```

@@

# events <- df$events %>%
#   mutate(src = factor(src,
#                       levels = 0:(nrow(demes) - 1),
#                       #labels = sort(demes$deme)),
#                       labels = c("France", "Germany","China","Italy","OtherEuropean","Spain")),
#          dest = ifelse(is.na(dest), as.character(src),
#                        as.character(factor(dest,
#                                            levels = 0:(nrow(demes) - 1),
#                                            #labels = sort(demes$deme),
#                                            labels = c("France" , "Germany","China","Italy","OtherEuropean","Spain")))),
#          date_model = date(date_decimal(decimal_date(ymd(MRS)) - age)))

# Do this in a function, useful also for Claire
events <- df$events %>%
  mutate(src = factor(src,
                      levels = 0:(nrow(demes) - 1),
                      #labels = sort(demes$deme)),
                      labels = c("France", "Germany","China","Italy","OtherEuropean","Spain")),
         dest = (factor(dest,
                           levels = 0:(nrow(demes) - 1),
                           #labels = sort(demes$deme),
                           labels = c("France" , "Germany","China","Italy","OtherEuropean","Spain"))))
events[events$event == "O", "src"] = "China"
events[events$event == "O", "mult"] =  1

df_traj0 <- events %>%
  mutate(date = date(date_decimal(decimal_date(mrs) - age))) %>%
  select(-time, -age) %>%
  group_by_at(vars(-mult)) %>%
  summarise(N = sum(mult), .groups = "drop") %>%
  mutate(srctmp = src, destmp = dest) %>%
  pivot_longer(src:dest, names_to = "role", values_to = "deme") %>%
  mutate(event = ifelse(event == "M" & role == "src", "OM",
                        ifelse(event == "M" & role == "dest", "IM", event)),
         partner = ifelse(event == "OM", as.character(destmp),
                        ifelse(event == "IM", as.character(srctmp), NA))) %>% 
  select(-c("role", "srctmp", "destmp")) %>%
  filter(!is.na(deme)) %>%
  group_by_at(vars(-N)) %>%
  summarise(value = sum(N), .groups = "drop") %>%
  group_by(traj, deme, partner, event) %>%
  arrange(traj, date, deme) %>%
  mutate(cumvalue = cumsum(value)) %>%
  rename(var = event)

mig_cum = df_traj0 %>%
  filter(var %in% c("IM", "OM")) %>%
  group_by(traj, var, date, deme) %>%
  summarise(value = sum(value), .groups = "drop") %>%
  group_by(traj, var, deme) %>%
  arrange(traj, date, deme) %>%
  mutate(cumvalue = cumsum(value),
         partner = NA)

inout_pop = df_traj0 %>% 
  mutate(var = ifelse(var %in% c("O", "B", "IM"), "in_pop", "out_pop")) %>%
  group_by(traj, deme, var, date) %>%
  summarise(value = sum(value), .groups = "drop_last") %>%
  arrange(traj, date, deme) %>%
  mutate(cumvalue = cumsum(value))

active_pop = inout_pop %>%
  select(-cumvalue) %>%
  pivot_wider(names_from = var, values_from = value, values_fill = list(value = 0)) %>%
  mutate(active_pop = in_pop - out_pop) %>%
  pivot_longer(active_pop, names_to = "var", values_to = "value") %>%
  group_by(traj, deme, var) %>%
  arrange(traj, date, deme) %>%
  mutate(cumvalue = cumsum(value)) %>%
  select(-c("in_pop", "out_pop"))

df_traj <- bind_rows(df_traj0, mig_cum) %>%
  bind_rows(inout_pop) %>%
  bind_rows(active_pop)

df_traj_cases <- df_traj %>%
  mutate(traj = as.character(traj)) %>%
  bind_rows(case_data %>%
            filter(cumvalue != 0) %>%
            mutate(var = "ECDC"))
                  # traj = "ecdc"))
```

@@

```{r traj-process, warning=FALSE}
# # Date grids
# max_age <- max(states$age)
# min_date <- date_decimal(decimal_date(mrs) - max_age)
# #ages_5day <- rev(decimal_date(mrs) - decimal_date(seq(mrs,min_date, by = '-5 days')))
# ages_1day <- rev(decimal_date(mrs) - decimal_date(seq(mrs,min_date, by = '-1 days')))
# 
# # Grid trajectories, time grid of 5 days taking maximum value
# gt <- gridTrajectoriesByAge(states, ages_1day) %>%
#   replace_na(list(N = 0)) %>%
#   mutate(date_model = date(date_decimal(decimal_date(mrs) - age)),
#          date = date_model + 10)
# 
# # Summary statistics
# gt_summary <- gt %>%
#   group_by(date, type) %>%
#   summarize(Imean = mean(N),
#             Imedian = median(N),
#             Ilow    = quantile(N, 0.025), #95% CI, compute HPD??
#             Ihigh   = quantile(N, 0.975),
#             .groups = "drop_last") %>%
#   ungroup()
# 
# # Grid events, time grid of 1 day, taking the sum of events
# ge <- gridEventsByAge(events, ages_1day) %>%
#   mutate(date_model = date(date_decimal(decimal_date(mrs) - age)),
#          date = date_model + 10,
#          dest = ifelse(is.na(dest), as.character(src), as.character(dest))) %>%
#   ungroup %>%
#   filter(N != 0)

@@

```{r}
# Active cases
d = "France"
df_traj_cases %>%
    # 1. Prepare data
    filter(deme == d, var %in% c("active_pop", "in_pop", "ECDC"), is.na(partner)) %>%
    # 2. Plot
    ggplot() +
    geom_line(aes(date, cumvalue, group = traj,
                  color = var)) +
                  # color = traj == "ecdc",
                  # alpha = traj == "ecdc",
                  # linetype = traj == "ecdc",
                  # size = traj == "ecdc")) +
    # scale_color_manual(name = "",
    #                    labels = c(paste0(deme, " inferred\n population trajectories"),
    #                               paste0(ifelse(deme == china_deme, "China", as.character(deme)), 
    #                                      " ECDC total case count")),
    #                    values = c(dcolors[[deme]], dark_dcolors[[deme]])) +
    # scale_alpha_manual(name = "", values = c(0.2, 1), guide = FALSE) +
    # scale_linetype_manual(name = "", values = c(1, 2),
    #                       labels = c(paste0(deme, " inferred\n population trajectories"),
    #                                  paste0(ifelse(deme == china_deme, "China", as.character(deme)), 
    #                                         " ECDC total case count"))) +
    # scale_size_manual(name = "", values = c(0.5, 1),
    #                   labels = c(paste0(deme, " inferred\n population trajectories"),
    #                              paste0(ifelse(deme == china_deme, "China", as.character(deme)), 
    #                                     " ECDC total case count"))) +
    scale_x_date(limits = c(ymd("2019-11-01"), ymd(MRS + 10)), date_breaks = "2 weeks", 
                 date_labels = "%b %d") +
    scale_y_log10(expand = c(0, 0),labels = trans_format("log10", math_format(10^.x))) +
    ylab("Population size") +
    theme(legend.position = c(0.4, 0.9),
          axis.title.x=element_blank())
```
